### 2019-01-30 LaTex/Paper_fulltext.tex
corrected gitignore in tex

%% 
%% Copyright 2007-2018 Elsevier Ltd
%%

\documentclass[preprint,12pt,authoryear]{elsarticle}

\usepackage{blindtext}
%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
%%\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

\journal{Dr. Florian Bader}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \address[label1]{}
%% \address[label2]{}

\title{Visualising Life Satisfaction in Countries across Europe: Individual- and Multi-Level Evidence on Cultural and Pecuniary Factors}

\author{Nicolas F. K. Göller\fnref{label2}}
\ead{n.goeller@zeppelin-university.net}
\fntext[label2]{Student in Sociology, Politics and Economics}

\author{Noah Peters\fnref{label1}}
\ead{no.peters@zeppelin-university.net}
\fntext[label2]{Student in Sociology, Politics and Economics}

\address{Zeppelin University, Fallenbrunnen 3, 88045 Friedrichshafen\fnref{label3}}
%% \fntext[label3]{}

\begin{abstract}
%% Text of abstract
This project was set out to provide an easy access to a comprehensible way for exploring the differences in determinants of life satisfaction 
all over Europe. While the actual analysis and presentation of results will be shown in different ways, this document
shall provide an overview over the data used and the treatment of variables applied during the project. 
The project's aim is not only to inspire interest into the determinants of happiness in human life, but also to encourage
later researchers to critically enquire with the work done here in order to point out its shortcomings and improve the final
result. Thus this document should provide the necessary transparency about sourcing of data and work on variables that
has been done previously to conducting the analysis and visualising results.
\end{abstract}

\end{frontmatter}

%% main text
\newpage
\tableofcontents
\newpage

\section{Introduction and Preliminary remarks}
Anecdotal evidence suggests that people spend a significant share of their lifetime working (for a casual calculation 
see: \citet{thompson_what_2016}). Moreover, people's work environment is changing constantly and especially today's 
discussion about the impact of digitalisation on the labour market requires to investigate how occupational factors shape 
people's life in general. Since work constitutes a considerable part of people’s life, satisfaction with one’s job and the
occupational environment must severely determine overall well-being. Drawing on the above-mentioned diversity of 
occupations, their respective circumstances and adaptation to the ‘digital revolution’, this assumption gains even more 
ground and relevance: \textit{If work changes, life changes.} \\
But which work-related factors contribute to a happy - or unhappy - life? Also, how does life satisfaction of the employed
vary across regions and countries? \\
Using data form the European Values Survey (EVS), we calculated multiple regression models to assess the causality of
work-related factors and well-being on the individual level. Moreover, multi-level analyses with regard to social capital and
macroeconomic as well as developmental indicators were performed to account for variation across regions and countries.

\section{Theory and Literature Review}
Although the causal relation between work and happiness/life satisfaction seems straightforward and intuitively correct,
its academic investigation is far from trivial. On the one hand, this appears in light of definitional questions regarding the 
wo variables of interest. On the other hand, operationalising the constructs appropriately remains subject to discussions
(for subjective well-being see: \cite{kahneman_developments_2006,layard_measuring_2010}). Lastly, the causal nature of job and life
satisfaction presents itself as far more complex than intuitively assumed. \\
Each of these three fields is addressed by a comprehensive literature review to set the tone for the subsequent analysis.
As this analysis draws on specific independent variables on the individual level and several factors on higher aggregate
levels (multi-level analysis), their respective relation to life satisfaction, as assessed in the literature, is discussed as well.

\subsection{Life Satisfaction as a Component of Subjective Well-Being (SWB)}
So far, the terms life satisfaction, happiness, and well-being have been used interchangeably. In fact, however, they
are commonly understood to act upon each other therefore constituting the wider concept of subjective well-being (SWB)
that includes life satisfaction, happiness, and positive affect  \citep{bowling_meta-analytic_2010,diener_subjective_1984}. 
Paying debt to the European Value Study’s (EVS) operationalisation of SWB, we only include life satisfaction as a
dependent variable: While life satisfaction is measured using a ten-point scale the other available component of SWB,
happiness, comes in a four-point scale. Consequently, we only included the life satisfaction variable for our regression
models. For the remains of this study, we concentrate on life satisfaction as a dependent variable. The variable selection
is further elaborated upon in the method section.

\subsection{SWB and Job Satisfaction}
The literature examining the causal relation of SWB and job satisfaction is extensive
(for reviews see: \citet{bowling_meta-analytic_2010,erdogan_whistle_2012}). \citet{bowling_meta-analytic_2010}, in their
meta-analysis, show a positive relationship between all three respective components of SWB and job satisfaction.
Going even further, \citet{near_job_1987} postulate that “job satisfaction must logically constitute some part of the
broader construct of life satisfaction; thus, the question regarding spillover between the two must be ‘how much’ not
‘whether’." (p. 398). \\
Before we examine the how much-question, it is necessary to discuss the direction of causality between SWB and job
satisfaction. In general, the present literature differentiates two approaches: the part-whole theory
\citep{bowling_meta-analytic_2010,hart_predicting_1999,judge_dispositional_1998} and the dispositional approach. The former
proposes job satisfaction to act upon SWB. This appears because of the simple reason that has already been mentioned
above: job satisfaction constitutes a sub-category of overall life satisfaction \citep[cf.]{near_job_1987}.
Consequently, the approach is also conceived of as bottom-up approach \citep{erdogan_whistle_2012}. Conversely, the
dispositional approach assumes a general predisposition of satisfaction with life to affect more specific domains like work
i.e., satisfaction with those domains \citep{bowling_meta-analytic_2010,diener_subjective_1984,headey_top-down_1991,judge_job_1993,judge_another_1993}. 
Erdogan et al. (2012) link these predispositions to personal traits that determine how satisfied people feel with certain
domains and life in general (top-down approach). Positioned between the poles is the spillover hypothesis that just
contents the basic linkage between job and life satisfaction or even other domains of life 
\citep{bowling_meta-analytic_2010,erdogan_whistle_2012,judge_job_2001,judge_individual_1994}.
Conceptually, this relation can be understood as a reciprocal relation between the various domains of life and people’s
satisfaction (Bowling et al., 2010). \\
While the direction of causality between life satisfaction and job satisfaction is an open discussion, the interplay of life
satisfaction with other work-related variables might be easier to grasp. Indeed, our study also examines the effects of
variables that characterise the working environment beyond perceived job satisfaction. This wider scope appears
reasonable since “job satisfaction is only a partial indicator of satisfaction with work domain. A person’s satisfaction with
career and work-related stress are important elements of the work domain that also need to be understood.” \citep[p. 1042]{erdogan_whistle_2012}.

%% The Appendices part is started with the command \appendix; appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% If you have bibdatabase file and want bibtex to generate the bibitems, please use
\bibliographystyle{elsarticle-harv} 
\bibliography{Literature}

\end{document}

\endinput
%%

### 2019-01-30 LaTex/Paper_fulltext.tex
Manifesto first draft
cdbk.tex

@@
 vary across regions and countries? \\
-Using data form the European Values Survey (EVS), we calculated multiple regression models to assess the causality of
+Using data form the European Values Survey (EVS) and several other sources, we calculated multiple regression models to assess the causality of
 work-related factors and well-being on the individual level. Moreover, multi-level analyses with regard to social capital and
 macroeconomic as well as developmental indicators were performed to account for variation across regions and countries. 
+%% Könntest du den Fokus ein wenig weg von Work und hin auf multilevel + pecuniary/ non-pecuiary setzen, für so viel Fokus haben wir nicht genug work-relatedness
+\section{Motivational Manifesto}
+In a time of widespread digitalisation for many aspects of modern society, 'datafication' and subsequent analysis of data has become ever more prevalent \cite{baack_datafication_2015, lycett_datafication:_2013}.
+Transparency of data sourcing, data access and methods employed in the course of this analysis is a chief concern of the 'Open Science' movement\cite{miguel_promoting_2014}.
+With more and more transparency, reproducibility of procedures is expected to increase and thus the quality of insights by replication and and improvements of methods through a 'wisdom of the (educated) crowd'.
+There have been extensive developments in the form of guidelines and protocols for researchers, universities and scientific journals that are aimed at promoting more transparency in the scientific process \cite{nosek_promoting_2015}.
+At the same time, large shares of the population lack the essential skills to participate in the whole process or perceive the barrier of entry in acquiring the necessary skills to do so, as unsurmountably high.
+The concept of 'Citizen Science' has thus emerged sitting at the edge of both science and education.
+Citizen science projects enlist non-scientiest members of the public to participate in the research (mainly in data collection), allowing to survey large amounts of data, while educating participants about the research topic \cite{shirk_public_2012}.
+Originally coming from environmental research this approach has already allowed researcher to survey wide ranges of locations and time periods and thus provides access to high-quality data that could quite impossibly be attained by professional research teams on their own \cite{bonney_citizen_2009}.
+
+Conceptually, 'Open Science' and 'Citizen Science' are deeply intertwined as 'Citizen Science' needs to be as transparent and easily accessible as possible to be effective on a large scale and 'Open Science' needs to offer educational value and low entry barriers if it wants to gain attention and participation of a broader audience.
+We feel committed to the mission of these movements and have thus designed our project in order to advance it for the social sciences.
 

### 2019-01-30 D_Visualisation.R

@@ -532,5 +532,5 @@ EVS_book <- within(EVS_book,{
 codebook(EVS_book)
 
 Write(codebook(EVS_book), file = "EVS_final_cdbk.txt")
-
+#Write(codebook(EVS_book), file = "LaTex/EVS_final_cdbk.tex")
 

### 2019-01-30 LaTex/Paper_fulltext.tex
Final workings on app, codebook; Problems with TEX

@@
 In a time of widespread digitalisation for many aspects of modern society, 'datafication' and subsequent analysis of data has become ever more prevalent \cite{baack_datafication_2015, lycett_datafication:_2013}.
-Transparency of data sourcing, data access and methods employed in the course of this analysis is a chief concern of the 'Open Science' movement\cite{miguel_promoting_2014}.
+Transparency of data sourcing, data access and methods employed in the course of this analysis is a chief concern of the 'Open Science' movement\cite{delfanti_open_2010}.
 With more and more transparency, reproducibility of procedures is expected to increase and thus the quality of insights by replication and and improvements of methods through a 'wisdom of the (educated) crowd'.
-There have been extensive developments in the form of guidelines and protocols for researchers, universities and scientific journals that are aimed at promoting more transparency in the scientific process \cite{nosek_promoting_2015}.
-At the same time, large shares of the population lack the essential skills to participate in the whole process or perceive the barrier of entry in acquiring the necessary skills to do so, as unsurmountably high.
+There have been extensive developments in the form of guidelines and protocols for researchers, universities and scientific journals that are aimed at promoting more transparency in the scientific process \cite{nosek_promoting_2015, miguel_promoting_2014}.
+At the same time, large shares of the population lack the essential skills to participate in the research process or perceive the barrier of entry in acquiring the necessary skills to do so, as unsurmountably high.
 The concept of 'Citizen Science' has thus emerged sitting at the edge of both science and education.
@@
 
-Conceptually, 'Open Science' and 'Citizen Science' are deeply intertwined as 'Citizen Science' needs to be as transparent and easily accessible as possible to be effective on a large scale and 'Open Science' needs to offer educational value and low entry barriers if it wants to gain attention and participation of a broader audience.
+Conceptually, 'Open Science' and 'Citizen Science' are deeply intertwined as 'Citizen Science' needs to be as transparent and easily accessible as possible to be effective on a large scale\cite{nov_dusting_2011} and 'Open Science' needs to offer educational value and low entry barriers if it wants to gain attention and participation of a broader audience \cite{newman_future_2012}.
 We feel committed to the mission of these movements and have thus designed our project in order to advance it for the social sciences.

### 2019-01-30 A_Data preparation.R

@@ -21,11 +21,11 @@ EVS <- EVS_2008 %>%
   select( #Pre-selection of variables
     country, c_abrv, v371b_N1, v66, v8, #base
     v62, v63, v90, v339SIOPS, #non-pecuniary
-    v89, v353YR, v353MM, v353M_ppp, v339ISEI, v198, #pecuniary
+    v89, v353_r, v353YR, v353MM, v353M_ppp, v339ISEI, v198, #pecuniary
     v1, v91, v92, v95, v96, v341, v338, #work variables
     v205:v218, v222, #index institutional trust
     v233, v234, v235, v237, v239, v245, v247, #index justification
-    v302, v303, v313, v336, v336_2, v370a) #demographics
+    v302, v303, age_r2, v313, v336_r, v336, v336_2, v370a) #demographics
 
 #---------------------------------------#
 #Export new dataset EVS                 #

### 2019-01-30 B_Variable preparation.R

@@ -178,11 +178,11 @@ EVS %<>% within({ #non-pecuniary factors
   job_sat <- v90 #Job satisfaction (10 point)
   job_sat[v90 %in% c(-5, -4, -3, -2, -1)] <- NA
   job_sat <- as.numeric(job_sat)
-  job_sat_c <- job_sat - mean(job_sat, na.rm = T)
   
   siops <- v339SIOPS #Standard Index of Occupational Prestige Scala (1-100)
+  siops[v339SIOPS %in% c(-5, -4, -3, -2, -1)] <- NA 
   siops <- as.numeric(siops)
-  siops_c <- siops - mean(siops, na.rm = T)
+  
   
 }) #non-pecuniary factors
 
@@ -196,29 +196,29 @@ EVS %<>% within({ #pecuniary factors
   nowork <- factor(nowork, levels = c(0, 1), labels = c("Employed", "Unemployed"), 
                        ordered = F)
   
+  inc_cat <- v353_r #Income recoded in three categories
+  inc_cat[v353_r %in% c(-5, -4, -3, -2, -1)] <- NA
+  inc_cat <- factor(inc_cat, levels = c(1, 2, 3), labels = c("Low", "Medium", "High"))
+  
   inc_an <- v353YR #Annual income in euros
   inc_an[v353YR %in% c(-5, -4, -3, -2, -1)] <- NA
   inc_an <- as.numeric(inc_an)
-  inc_an_c <- inc_an - mean(inc_an, na.rm = T)
   
   inc_mon <- v353MM #Monthly income in euros
   inc_mon[v353MM %in% c(-5, -4, -3, -2, -1)] <- NA
   inc_mon <- as.numeric(inc_mon)
-  inc_mon_c <- inc_mon - mean(inc_mon, na.rm = T)
   
   incppp_mon <- v353M_ppp #Monthly income after purchasing power parity
   incppp_mon[v353M_ppp %in% c(-5, -4, -3, -2, -1)] <- NA
   incppp_mon <- as.numeric(incppp_mon)
-  incppp_mon_c <- incppp_mon - mean(incppp_mon, na.rm = T)
   
   isei <- v339ISEI #International Socio-Economic Index of Occupational Status
+  isei[v339ISEI %in% c(-5, -4, -3, -2, -1)] <- NA 
   isei <- as.numeric(isei)
-  isei_c <- isei - mean(isei, na.rm = T)
   
[…]

### 2019-01-30 D_Visualisation.R

@@ -171,8 +171,8 @@ EVS_book <- dplyr::select(EVS_final,
                           cl_sb, ch_tax, joy, lying, bribe, av_tax, av_pub_f,  
                           trust_wrth,
                           work_impo, free_job, talent, duty, work_first, sup, entre,
-                          isei, inc_mon, inc_an, incppp_mon, inc_eq,
-                          sex, age, mar_stat, edu_cat, edu, town,
+                          isei, inc_mon, inc_cat, inc_an, incppp_mon, inc_eq,
+                          sex, age, age_cat, mar_stat, edu_cat3, edu_cat7, edu, town,
                           fhrate, hdi, gini, GDPpc, unemployment, life_sat,
                           intp_trust_reg, inst_trust_reg, trust_wrth_reg, fair_reg)
 
@@ -403,12 +403,11 @@ EVS_book <- within(EVS_book,{
   annotation(isei)["Source"] <- "European Value Survey 2008 (v339ISEI)"
   annotation(isei)["Remark"] <- "The ISEI-Index is  a widely known measure for socio-economic status of different occupations."
   
-  description(inc_mon) <- "Individual level: Monthly income"
-  wording(inc_mon) <- "Here is a list of incomes and we would like to know in what group your household is, counting all wages, salaries, pensions and other incomes that come in. Just give the letter of the group your household falls into, after taxes and other deductions. (12 categories)"
-  measurement(inc_mon) <- "interval"
-  missing.values(inc_mon) <- c(NA)
-  annotation(inc_mon)["Source"] <- "European Value Survey 2008 (v353MM)"
-  annotation(inc_mon)["Remark"] <- "The assumption of quasi-metric variable was taken when recoding to numeric."
+  description(inc_cat) <- "Individual level: Income recoded categorical"
+  measurement(inc_cat) <- "ordinal"
+  missing.values(inc_cat) <- c(NA)
+  annotation(inc_cat)["Source"] <- "European Value Survey 2008 (v353_r)"
+  annotation(inc_cat)["Remark"] <- "The variable taken as given, recoded by the data publisher. Transformed to ordered factor."
   
   description(inc_an) <- "Individual level: Annual income"
[…]

### 2019-01-30 European_Happiness_Observer/app.R

@@ -8,6 +8,7 @@ library(magrittr)
 library(hrbrthemes)
 library(lme4)
 library(stargazer)
+library(viridis)
 
 #------------------------------------------------------------------------------------
 
@@ -19,8 +20,8 @@ shiny_data <- dplyr::select(EVS_final,
                             intp_trust,inst_trust, trust_wrth,
                             work_impo, free_job, talent, duty, work_first, sup, 
                             entre, nowork,
-                            isei, inc_mon, inc_an, incppp_mon, inc_eq,
-                            sex, age, mar_stat, edu_cat, edu, town, 
+                            isei, inc_cat, inc_an, inc_mon, incppp_mon, inc_eq,
+                            sex, age, age_cat, mar_stat, edu_cat3, edu_cat7, edu, town, 
                             nation, eureg, reg, 
                             hdi, gini, fhrate, unemployment, GDPpc, life_sat,
                             trust_wrth_reg, intp_trust_reg, inst_trust_reg, fair_reg)
@@ -45,15 +46,18 @@ shiny_data %<>% within({
   Unemployed <- nowork
   
   ISEI_Index <- isei
-  Monthly_income <- inc_mon
   Annual_income <- inc_an
+  Income_categories <- inc_cat
+  Monthly_income <- inc_mon
   Monthly_income_ppp <- incppp_mon
   Equality_of_income_incentives <- inc_eq
   
   Sex <- sex
   Age <- age
-  Education_categories <- edu_cat
-  Education <- edu
+  Age_categories <- age_cat
+  Education_categories <- edu_cat3
+  Education <- edu_cat7
+  Education_scale <- edu
   Marital_status <- mar_stat
   Town_size <- town
   
@@ -73,26 +77,28 @@ shiny_data %<>% within({
   Regional_interpersonal_trust <- intp_trust_reg
   Regional_expected_fairness <- fair_reg
 })
-shiny_data <- dplyr::select(shiny_data, #from now: 1:27
-                            Life_satisfaction, Happiness, SIOPS_Index, Job_satisfaction, 
-                            Fairness_of_other_people, Institutional_trust, Trustworthiness, 
-                            Trust_interpersonal, Importance_of_work, Freedom_in_job_decisions,
-                            Job_talent_development, Work_as_societal_duty, Work_devotion,
-                            Supervisory_work, Entrepreneurial_work, Unemployed, 
-                            ISEI_Index, Monthly_income, Annual_income, Monthly_income_ppp,
[…]

### 2019-01-30 LaTex/Paper_fulltext.tex
Added 'Data' part
Correct European Value Survey to Study where present

--- 2019-01-30 LaTex/Paper_fulltext.tex
@@
 
-\usepackage{blindtext}
 %% For including figures, graphicx.sty has been loaded in
@@
 	vary across regions and countries? \\
-Using data form the European Values Survey (EVS), we calculated multiple regression models to assess the causality of
+	Using data form the European Values Study (EVS) and several other sources, we calculated multiple regression models to assess the causality of
 	work-related factors and well-being on the individual level. Moreover, multi-level analyses with regard to social capital and
 	macroeconomic as well as developmental indicators were performed to account for variation across regions and countries. 
+	%% Könntest du den Fokus ein wenig weg von Work und hin auf multilevel + pecuniary/ non-pecuiary setzen, für so viel Fokus haben wir nicht genug work-relatedness
+	\section{Motivational Manifesto}
+	In a time of widespread digitalisation for many aspects of modern society, 'datafication' and subsequent analysis of data has become ever more prevalent \citep{baack_datafication_2015, lycett_datafication:_2013}.
+Transparency of data sourcing, data access and methods employed in the course of this analysis is a chief concern of the 'Open Science' movement \citep{delfanti_open_2010}.
+With more and more transparency, reproducibility of procedures is expected to increase and thus the quality of insights by replication and and improvements of methods through a 'wisdom of the (educated) crowd'.
+There have been extensive developments in the form of guidelines and protocols for researchers, universities and scientific journals that are aimed at promoting more transparency in the scientific process \citep{nosek_promoting_2015, miguel_promoting_2014}.
+	At the same time, large shares of the population lack the essential skills to participate in the research process or perceive the barrier of entry in acquiring the necessary skills to do so, as unsurmountably high.
+The concept of 'Citizen Science' has thus emerged sitting at the edge of both science and education.
+Citizen science projects enlist non-scientiest members of the public to participate in the research (mainly in data collection), allowing to survey large amounts of data, while educating participants about the research topic \citep{shirk_public_2012}.
+Originally coming from environmental research this approach has already allowed researcher to survey wide ranges of locations and time periods and thus provides access to high-quality data that could quite impossibly be attained by professional research teams on their own \citep{bonney_citizen_2009}.
+
+	Conceptually, 'Open Science' and 'Citizen Science' are deeply intertwined as 'Citizen Science' needs to be as transparent and easily accessible as possible to be effective on a large scale \citep{nov_dusting_2011} and 'Open Science' needs to offer educational value and low entry barriers if it wants to gain attention and participation of a broader audience \citep{newman_future_2012}.
+We feel committed to the mission of these movements and have thus designed our project in order to advance it for the social sciences.
 
@@
 	wo variables of interest. On the other hand, operationalising the constructs appropriately remains subject to discussions
-(for subjective well-being see: \cite{kahneman_developments_2006,layard_measuring_2010}). Lastly, the causal nature of job and life
+	(for subjective well-being see: \citet{kahneman_developments_2006,layard_measuring_2010}). Lastly, the causal nature of job and life
 	satisfaction presents itself as far more complex than intuitively assumed. \\
@@
 
+	\section{Data}
+	All data employed in this project i acquired from official sources and with permission of the provider or in the course of open access.
+Data as well as original codebooks will be made available for the user upon request, or delivered within the project.
+	The project is mainly built on data from the 'European Value Study 2008' (EVS), which is the fourth survey wave \citep{evs_european_2016}.
+The EVS contains questions regarding values of Europeans on aspects like moral, work, family, politics, society, religion, as well as an extensive collection of sociodemographic items.
+Issues and questions regarding surveying of the data, errata, harmonizing of results etc.
+may be settled with the publisher or from the original codebook.
+	The file contains a total 66281 observations for 477 variables over 46 European countries.
+All transformations on variables can be read in detail in the codebook.
+Significant changes to the data were made through exclusion of 'Russian Federation' from the sample for the purpose of better visualisation.
+Furthermore, 'Northern Cyprus' and 'Northern Ireland' which were coded as own countries were added to their respective states after international law.
+Thus, data for 43 countries was processed for use in the project.
+Furthermore, aggregations were made from the data in total of 5 cases to create variables that are applicable for multilevel analysis.
+Finally, a variable of 'Geographical region' was created from the country variable in the EVS.
+This classification was based on the 'Standard Country or Area Codes for Statistics Use' from the UN Statistics Division.
+	Structural data for nations was obtained from the World Bank for 'GDP per capita of 2008' (in 2018 dollars per inhabitant), and for 'Gini Coefficient 2008' (in ration scale 0-1).
+In the case of 'Gini Coefficient 2008', missing values for Germany, Poland, Serbia, Kosovo, Macedonia, Croatia and Bosnia Herzegovina were taken from Eurostat 2008 (Germany, Poland), from the CIA World Factbook (Serbia 2008) from World Bank Gini Coefficient 2009 (Kosovo, Macedonia, Croatia), and from World Bank Gini Coefficient 2007 (Bosnia Herzegovina).
+Data was also obtained from the 'Democracy Cross-National' dataset created by Pippa Norris for the Freedom House Democracy rating 2008 and the Human Development Index 2008.
+Missing values of the Freedom House rating for Kosovo and Montenegro were replaced by the indication for 2008 on the Freedom House website.
+Missing values of the HDI for Kosovo was approximated by the value for Serbia in 2008.
+
+\section{Responsibility for Project Areas}
+
+	\begin{tabular}{|c|c|c|}
+		\hline 
+		\textbf{Project area}	& \textbf{Noah Peters} 	& \textbf{Nicolas Goeller} 	\\ 
+		\hline 
+		Data wrangling			&  		   				&  				 			\\ 
+		\hline 
+		Variable preparation 	&  		   				&  				 			\\ 
+		\hline 
+		Geodata					&  		   				&  				 			\\ 
+		\hline 
+		Git	     			 	&  		   				&  				 			\\ 
+		\hline 			
+		Indices				 	&  		   				&  				 			\\ 
+		\hline 
+		Paper visualisations 	&  		   				&  				 			\\ 
+		\hline 
+		Shiny 				 	&  		   				&  				 			\\ 
+		\hline 
+		Codebook 			 	&  		   				&  				 			\\ 
+		\hline 
+		Geoplots 			 	&  		   				&  				 			\\ 
+		\hline 
+		LaTex				 	&  		   				&  				 			\\ 
+		\hline 
+		Motivation			 	&  		   				&  				 			\\ 
+		\hline 
+		Theory				 	&  		   				&  				 			\\ 
+		\hline 
+		User manual			 	&  		   				&  				 			\\ 
+		\hline 
+	\end{tabular}
+
 	%% The Appendices part is started with the command \appendix; appendix sections are then done as normal sections

### 2019-01-30 A_Data preparation.R

@@ -11,7 +11,7 @@ library(magrittr)
 
 #-----------------------------------------------------------------------------------------
 
-###  1. European Value Survey 2008
+###  1. European Value Study 2008
 EVS_2008 <- read_spss("Data/ZA4800_v4-0-0.sav") # Seems to have the missing countries
 
 #Preliminary exclusions and selection

### 2019-01-30 B_Variable preparation.R

@@ -20,9 +20,9 @@ library(eurostat)
 
 #-------------------------------------------------------------------------------------------------------------------------------------------------------------------
 #-------------------------------------------------------------------------------------------------------------------------------------------------------------------
-###   1. European Values Survey
+###   1. European Values Study
 
-##    1.1 Read European Values Survey 2008
+##    1.1 Read European Values Study 2008
 EVS <- read_rds("Data/EVS.rds")
 
 
@@ -509,7 +509,7 @@ wb_data$nation <- wb_data$country
 
 #----------------------------------------------------------------------------------------
 
-##    2.4 European Value Survey 2008: Level-2-variables
+##    2.4 European Value Study 2008: Level-2-variables
 
 # Creating a subsetted dataset for variable to be aggregated
 EVS_nat <- dplyr::select(EVS, nation, nowork, sat) 

### 2019-01-30 D_Visualisation.R

@@ -177,80 +177,80 @@ EVS_book <- dplyr::select(EVS_final,
                           intp_trust_reg, inst_trust_reg, trust_wrth_reg, fair_reg)
 
 EVS_book <- as.data.set(EVS_book)
-  
+  n_distinct(EVS_final$country)
 EVS_book <- within(EVS_book,{
   
   description(nation) <- "Country of residence"
   measurement(nation) <- "nominal"
   missing.values(nation) <- c(NA)
-  annotation(nation)["Source"] <- "European Value Survey 2008 (country)"
+  annotation(nation)["Source"] <- "European Value Study 2008 (country)"
   annotation(nation)["Remark"] <- "The Russian Federation was excluded for purposes of visualisation. Northern Ireland and Northern Cyprus were recoded to belong to their respective legal states. Azerbaijan was excluded as there were no respondents in the sample."
   
   description(c_code) <- "Country Code"
   measurement(c_code) <- "nominal"
   missing.values(c_code) <- c(NA)
-  annotation(c_code)["Source"] <- "European Value Survey 2008 (c_abrv)"
+  annotation(c_code)["Source"] <- "European Value Study 2008 (c_abrv)"
   annotation(c_code)["Remark"] <- "Country codes were recoded to match changes in country selection (see 'Country of residence)."
   
   description(reg) <- "Region of residence"
   measurement(reg) <- "nominal"
   missing.values(reg) <- c(NA)
-  annotation(reg)["Source"] <- "European Value Survey 2008 (v371b_N1)"
+  annotation(reg)["Source"] <- "European Value Study 2008 (v371b_N1)"
   annotation(reg)["Remark"] <- "Based on NUTS-1 coding, region codes were recoded to match changes in country selection."
   
   description(eureg) <- "Geopgraphical region in Europe"
   measurement(eureg) <- "nominal"
   missing.values(eureg) <- c(NA)
-  annotation(eureg)["Source"] <- "European Value Survey 2008, UN geograhical region code"
+  annotation(eureg)["Source"] <- "European Value Study 2008, UN geograhical region code"
   annotation(eureg)["Remark"] <- "The item was recoded after the UN classification from the variable 'nation'. 'Armenia' was added to Southern Europe and 'Georgia' was added to Eastern Europe."
   
   description(sat) <- "Individual level: Life satisfaction"
[…]

### 2019-01-30 European_Happiness_Observer/app.R

@@ -234,8 +234,6 @@ ui <- dashboardPage(
                     width = 3,
                     "Choose your variable for plotting",br(), br(),
                     varSelectInput("macro", "Variable:", shiny_nat[,1:7]))
-                #box(title = "Data indicator", status = "success", solidHeader = T, width = 3,
-                #    selectInput("nat", choices = nat))
               )),
       tabItem(tabName = "regmap", h1("Geographical overview"),
               fluidRow(

### 2019-01-31 LaTex/Paper_fulltext.tex
Added citations, solved warnings and problems

@@
 %% \usepackage{amsthm}
-
+\usepackage{url}
+\usepackage{natbib}
 %% The lineno packages adds line numbers. Start line numbering with
@@
 Finally, a variable of 'Geographical region' was created from the country variable in the EVS.
-This classification was based on the 'Standard Country or Area Codes for Statistics Use' from the UN Statistics Division.
-	Structural data for nations was obtained from the World Bank for 'GDP per capita of 2008' (in 2018 dollars per inhabitant), and for 'Gini Coefficient 2008' (in ration scale 0-1).
-In the case of 'Gini Coefficient 2008', missing values for Germany, Poland, Serbia, Kosovo, Macedonia, Croatia and Bosnia Herzegovina were taken from Eurostat 2008 (Germany, Poland), from the CIA World Factbook (Serbia 2008) from World Bank Gini Coefficient 2009 (Kosovo, Macedonia, Croatia), and from World Bank Gini Coefficient 2007 (Bosnia Herzegovina).
-Data was also obtained from the 'Democracy Cross-National' dataset created by Pippa Norris for the Freedom House Democracy rating 2008 and the Human Development Index 2008.
-Missing values of the Freedom House rating for Kosovo and Montenegro were replaced by the indication for 2008 on the Freedom House website.
-Missing values of the HDI for Kosovo was approximated by the value for Serbia in 2008.
+This classification was based on the 'Standard Country or Area Codes for Statistics Use' from the UN Statistics Division \citep{unsd_standard_2019}.
+	Structural data for nations was obtained from the World Bank for 'GDP per capita of 2008' (in 2018 dollars per inhabitant)\citep{world_bank_gdp_2019}, and for 'Gini Coefficient 2008' (in ration scale 0-1)\citep{world_bank_gini_2019}.
+In the case of 'Gini Coefficient 2008', missing values for Germany, Poland, Serbia, Kosovo, Macedonia, Croatia and Bosnia Herzegovina were taken from Eurostat 2008 (Germany, Poland) \citep{eurostat_eurostat_2019}, from the CIA World Factbook (Serbia 2008) \citep{central_intelligence_agency_serbia_2019}, from World Bank Gini Coefficient 2009 (Kosovo, Macedonia, Croatia), and from World Bank Gini Coefficient 2007 (Bosnia Herzegovina) \citep{world_bank_gini_2019}.
+Data was also obtained from the 'Democracy Cross-National' dataset created by Pippa Norris \citep{norris_data_2015} for the Freedom House Democracy rating 2008 and the Human Development Index 2008.
+Missing values of the Freedom House rating for Kosovo \citep{freedom_house_kosovo_2012} and Montenegro \citep{freedom_house_montenegro_2012} were replaced by the indication for 2008 on the Freedom House website.
+The missing value of the HDI 2008 for Kosovo was approximated by the value for Serbia in 2008.
+
+\section{User Manual}
 

### 2019-01-31 LaTex/Paper_fulltext.tex
Some Latex changes

@@
 There have been extensive developments in the form of guidelines and protocols for researchers, universities and scientific journals that are aimed at promoting more transparency in the scientific process \citep{nosek_promoting_2015, miguel_promoting_2014}.
-	At the same time, large shares of the population lack the essential skills to participate in the research process or perceive the barrier of entry in acquiring the necessary skills to do so, as unsurmountably high.
+	At the same time, large shares of the population lack the essential skills to participate in the research process or perceive the barrier of entry in acquiring the necessary skills to do so, as insurmountably high.
 The concept of 'Citizen Science' has thus emerged sitting at the edge of both science and education.
-Citizen science projects enlist non-scientiest members of the public to participate in the research (mainly in data collection), allowing to survey large amounts of data, while educating participants about the research topic \citep{shirk_public_2012}.
+Citizen science projects enlist non-scientist members of the public to participate in the research (mainly in data collection), allowing to survey large amounts of data, while educating participants about the research topic \citep{shirk_public_2012}.
 Originally coming from environmental research this approach has already allowed researcher to survey wide ranges of locations and time periods and thus provides access to high-quality data that could quite impossibly be attained by professional research teams on their own \citep{bonney_citizen_2009}.
@@
 	The project is mainly built on data from the 'European Value Study 2008' (EVS), which is the fourth survey wave \citep{evs_european_2016}.
-The EVS contains questions regarding values of Europeans on aspects like moral, work, family, politics, society, religion, as well as an extensive collection of sociodemographic items.
+The EVS contains questions regarding values of Europeans on aspects like moral, work, family, politics, society, religion, as well as an extensive collection of socio-demographic items.
 Issues and questions regarding surveying of the data, errata, harmonizing of results etc.
@@
 The file contains a total 66281 observations for 477 variables over 46 European countries.
-All transformations on variables can be read in detail in the codebook.
+All transformations on variables can be read in detail from the codebook.
 Significant changes to the data were made through exclusion of 'Russian Federation' from the sample for the purpose of better visualisation.
@@
 \section{User Manual}
+
+\section{Conclusion}
+
+	%% The Appendices part is started with the command \appendix; appendix sections are then done as normal sections
+\newpage
+ \appendix
 
@@
 
-	%% The Appendices part is started with the command \appendix; appendix sections are then done as normal sections
-	%% \appendix
-
-	%% \section{}
-	%% \label{}
-
 	%% If you have bibdatabase file and want bibtex to generate the bibitems, please use
+	\newpage
 	\bibliographystyle{elsarticle-harv} 

### 2019-01-31 LaTex/Paper_fulltext.tex

@@
 
+\subsection{The Wider Interplay of Work-Related Variables with Life Satisfaction}
+Our analysis features a myriad of variables that can be examined with regard to their effect on overall life satisfaction.
+In order to justify our variable selection the following remarks elucidate the respective variables' interplay with life satisfaction.
+In contrast to the previous examination of the overall concept of SWB, to accounts on the different variables are presented 
+in a briefer manner. Rather than covering the entire theoretical background, the following elaboration is though to spur interest
+and curiosity to examine the relation of life satisfaction and other factors on one's own - aided by our interactive 
+European Happiness Observer interface.
+
+\subsubsection{Life satisfaction and income}
+The life satisfaction-income discussion almost always refers back to \citet{easterlin_does_1974} whose observations regarding
+the relation of life satisfaction i.e., happiness and income were dubbed the `Easterlin Paradox'. Indeed, his findings from time series data
+suggest that life satisfaction remains flat, despite increases in GDP per capita. These findings still heavily inform other researchers'
+work and indicate that the relative income in comparision to others might influence SWB \citep{di_tella_happiness_2010,drichoutis_reference_2010,pedersen_happiness_2011}.
+Emphasising relative position follows the lead of behavioural labour economics that builds on the findings of prospect theory \citep{kahneman_prospect_1979}.
+This concept implies that reference points rather than absolute positions inform people's choices. In the course of behavioural
+labour economics, this backgroundis highly relevant \citep{dohmen_behavioral_2014}. 
+While investigations examining relative status and changes over time appear highly relevant and theoretically justified, the present
+project ist not suited to cover these questions. Not being in posession of either time series or relative income data, we need
+to focus on absolute income at a certain point in time. Although this independent variable might not capture the underlying relation
+of income and life satisfaction, it qualifies as a basic means of inquiry that sheds light on the distinction between life satisfaction
+and happiness: “high income buys life satisfaction but not happiness.” \citep[p. 16489]{kahneman_high_2010}.
+
+\subsubsection{Pecuniary vs. non-pecuniary factors}
+Income-related investigations also stimulate inquiries examining the opposite of monetary factors, namely non-pecuniary
+variables. \textit{Since high income might not be all to strive for, which `softer' factors determine a satisfactory life?}
+In general, health takes centre-stage as a determinant of life satisfaction. This statement is not only backed by common 
+sense or daily wisdom (Who has not said or been told that health is life's most precious gift?) but also by empirical observations \citep{pedersen_happiness_2011}. \\
+at the work place and terms like `work-life-balance' and `Generation Y' imply changing perceptions of a pleasant working environment. Also \citet{andersson_happiness_2008}
+investigates the special relation between self-employment and life satisfaction with ample regard to health issues. \\
+Self-employment as such exhibits a variable that is thought to affect life satisfaction \citep{andersson_happiness_2008,schneck_why_2014,van_der_zwan_self-employment_2018}
+due to non-pecuniary circumstances that are innate and to the status of self-employment. \\
+This remark also leads to a greater discussion of features proposed by self-determinantion theory \citep{deci_intrinsic_1985,deci_self-determination_1989,ryan_self-determination_2000}.
+Lastly, \citet{luo_essays_2018} uses EVS data estimate the effects of pecuniary and non-pecuniary factors on life satisfaction
+The EVS, for example, captures to which degree respondents feel the need for a job that enables employees to develop their talents. \\
+of the unemployed. Intuitively, financial uncertainty might appear as the major concern of unemployment but further investigations
+show the significant role that social and psychological factors play. This observation can be interpreted as a general plea for research on
+the interplay of non-pecuniary factors of the work environment and life satisfaction.
+
+\subsubsection{Non-pecuniary factors of the work environment}
+Drawing on the previous remarks regarding in incorporation of work-related non-pecuniary variables, our analysis comprises of
+multiple factors that touch on leadership responsibilities, self-determination related factors, or the occupation's reputation.
+The following variables can be examined using the European Happiness Observer:
+\begin{itemize}
+   \item Work importance in life
+   \item Freedem of job decisions
+   \item Job development
+   \item Work duty to society
+   \item Work should come always first
+   \item Supervising responsibilities 
+   \item Status of employment (i.e., being self-employed)
+   \item International Socio-Economic Index of Occupational Status   
+\end{itemize}
 %% If you have bibdatabase file and want bibtex to generate the bibitems, please use
-	\newpage
 \bibliographystyle{elsarticle-harv} 
 \bibliography{Literature}
-
+In the working environment, health is not a trivial factor either. The \citet{who_global_1994} stresses the importance of health
 \end{document}

### 2019-01-31 LaTex/Paper_fulltext.tex
Worked on problems with structure in the text but now everything is smooth

@@
 
+	\subsection{The Wider Interplay of Work-Related Variables with Life Satisfaction}
+Our analysis features a myriad of variables that can be examined with regard to their effect on overall life satisfaction.
+In order to justify our variable selection the following remarks elucidate the respective variables' interplay with life satisfaction.
+In contrast to the previous examination of the overall concept of SWB, to accounts on the different variables are presented 
+in a briefer manner. Rather than covering the entire theoretical background, the following elaboration is though to spur interest
+and curiosity to examine the relation of life satisfaction and other factors on one's own - aided by our interactive 
+European Happiness Observer interface.
+
+\subsubsection{Life satisfaction and income}
+The life satisfaction-income discussion almost always refers back to \citet{easterlin_does_1974} whose observations regarding
+the relation of life satisfaction i.e., happiness and income were dubbed the `Easterlin Paradox'. Indeed, his findings from time series data
+suggest that life satisfaction remains flat, despite increases in GDP per capita. These findings still heavily inform other researchers'
+work and indicate that the relative income in comparision to others might influence SWB \citep{di_tella_happiness_2010,drichoutis_reference_2010,pedersen_happiness_2011}.
+Emphasising relative position follows the lead of behavioural labour economics that builds on the findings of prospect theory \citep{kahneman_prospect_1979}.
+This concept implies that reference points rather than absolute positions inform people's choices. In the course of behavioural
+labour economics, this backgroundis highly relevant \citep{dohmen_behavioral_2014}. 
+While investigations examining relative status and changes over time appear highly relevant and theoretically justified, the present
+project ist not suited to cover these questions. Not being in posession of either time series or relative income data, we need
+to focus on absolute income at a certain point in time. Although this independent variable might not capture the underlying relation
+of income and life satisfaction, it qualifies as a basic means of inquiry that sheds light on the distinction between life satisfaction
+and happiness: “high income buys life satisfaction but not happiness.” \citep[p. 16489]{kahneman_high_2010}.
+
+\subsubsection{Pecuniary vs. non-pecuniary factors}
+Income-related investigations also stimulate inquiries examining the opposite of monetary factors, namely non-pecuniary
+variables. \textit{Since high income might not be all to strive for, which `softer' factors determine a satisfactory life?}
+In general, health takes centre-stage as a determinant of life satisfaction. This statement is not only backed by common 
+sense or daily wisdom (Who has not said or been told that health is life's most precious gift?) but also by empirical observations \citep{pedersen_happiness_2011}. \\
+In the working environment, health is not a trivial factor either.
+The \citet{who_global_1994} stresses the importance of health at the work place and terms like `work-life-balance' and `Generation Y' imply changing perceptions of a pleasant working environment.
+Also \citet{andersson_happiness_2008}
+investigates the special relation between self-employment and life satisfaction with ample regard to health issues. \\
+Self-employment as such exhibits a variable that is thought to affect life satisfaction \citep{andersson_happiness_2008,schneck_why_2014,van_der_zwan_self-employment_2018}
+due to non-pecuniary circumstances that are innate and to the status of self-employment. \\
+This remark also leads to a greater discussion of features proposed by self-determinantion theory \citep{deci_intrinsic_1985,deci_self-determination_1989,ryan_self-determination_2000}.
+Lastly, \citet{luo_essays_2018} uses EVS data estimate the effects of pecuniary and non-pecuniary factors on life satisfaction
+The EVS, for example, captures to which degree respondents feel the need for a job that enables employees to develop their talents. \\
+of the unemployed. Intuitively, financial uncertainty might appear as the major concern of unemployment but further investigations
+show the significant role that social and psychological factors play. This observation can be interpreted as a general plea for research on
+the interplay of non-pecuniary factors of the work environment and life satisfaction.
+
+\subsubsection{Non-pecuniary factors of the work environment}
+Drawing on the previous remarks regarding in incorporation of work-related non-pecuniary variables, our analysis comprises of
+multiple factors that touch on leadership responsibilities, self-determination related factors, or the occupation's reputation.
+The following variables can be examined using the European Happiness Observer:
+\begin{itemize}
+   \item Work importance in life
+   \item Freedem of job decisions
+   \item Job development
+   \item Work duty to society
+   \item Work should come always first
+   \item Supervising responsibilities 
+   \item Status of employment (i.e., being self-employed)
+   \item International Socio-Economic Index of Occupational Status   
+\end{itemize}
+
 \section{Data}
@@
 
-\subsection{The Wider Interplay of Work-Related Variables with Life Satisfaction}
-Our analysis features a myriad of variables that can be examined with regard to their effect on overall life satisfaction.
-In order to justify our variable selection the following remarks elucidate the respective variables' interplay with life satisfaction.
-In contrast to the previous examination of the overall concept of SWB, to accounts on the different variables are presented 
-in a briefer manner. Rather than covering the entire theoretical background, the following elaboration is though to spur interest
-and curiosity to examine the relation of life satisfaction and other factors on one's own - aided by our interactive 
-European Happiness Observer interface.
-
-\subsubsection{Life satisfaction and income}
-The life satisfaction-income discussion almost always refers back to \citet{easterlin_does_1974} whose observations regarding
-the relation of life satisfaction i.e., happiness and income were dubbed the `Easterlin Paradox'. Indeed, his findings from time series data
-suggest that life satisfaction remains flat, despite increases in GDP per capita. These findings still heavily inform other researchers'
-work and indicate that the relative income in comparision to others might influence SWB \citep{di_tella_happiness_2010,drichoutis_reference_2010,pedersen_happiness_2011}.
-Emphasising relative position follows the lead of behavioural labour economics that builds on the findings of prospect theory \citep{kahneman_prospect_1979}.
-This concept implies that reference points rather than absolute positions inform people's choices. In the course of behavioural
-labour economics, this backgroundis highly relevant \citep{dohmen_behavioral_2014}. 
-While investigations examining relative status and changes over time appear highly relevant and theoretically justified, the present
-project ist not suited to cover these questions. Not being in posession of either time series or relative income data, we need
-to focus on absolute income at a certain point in time. Although this independent variable might not capture the underlying relation
-of income and life satisfaction, it qualifies as a basic means of inquiry that sheds light on the distinction between life satisfaction
-and happiness: “high income buys life satisfaction but not happiness.” \citep[p. 16489]{kahneman_high_2010}.
-
-\subsubsection{Pecuniary vs. non-pecuniary factors}
-Income-related investigations also stimulate inquiries examining the opposite of monetary factors, namely non-pecuniary
-variables. \textit{Since high income might not be all to strive for, which `softer' factors determine a satisfactory life?}
-In general, health takes centre-stage as a determinant of life satisfaction. This statement is not only backed by common 
-sense or daily wisdom (Who has not said or been told that health is life's most precious gift?) but also by empirical observations \citep{pedersen_happiness_2011}. \\
-at the work place and terms like `work-life-balance' and `Generation Y' imply changing perceptions of a pleasant working environment. Also \citet{andersson_happiness_2008}
-investigates the special relation between self-employment and life satisfaction with ample regard to health issues. \\
-Self-employment as such exhibits a variable that is thought to affect life satisfaction \citep{andersson_happiness_2008,schneck_why_2014,van_der_zwan_self-employment_2018}
-due to non-pecuniary circumstances that are innate and to the status of self-employment. \\
-This remark also leads to a greater discussion of features proposed by self-determinantion theory \citep{deci_intrinsic_1985,deci_self-determination_1989,ryan_self-determination_2000}.
-Lastly, \citet{luo_essays_2018} uses EVS data estimate the effects of pecuniary and non-pecuniary factors on life satisfaction
-The EVS, for example, captures to which degree respondents feel the need for a job that enables employees to develop their talents. \\
-of the unemployed. Intuitively, financial uncertainty might appear as the major concern of unemployment but further investigations
-show the significant role that social and psychological factors play. This observation can be interpreted as a general plea for research on
-the interplay of non-pecuniary factors of the work environment and life satisfaction.
-
-\subsubsection{Non-pecuniary factors of the work environment}
-Drawing on the previous remarks regarding in incorporation of work-related non-pecuniary variables, our analysis comprises of
-multiple factors that touch on leadership responsibilities, self-determination related factors, or the occupation's reputation.
-The following variables can be examined using the European Happiness Observer:
-\begin{itemize}
-   \item Work importance in life
-   \item Freedem of job decisions
-   \item Job development
-   \item Work duty to society
-   \item Work should come always first
-   \item Supervising responsibilities 
-   \item Status of employment (i.e., being self-employed)
-   \item International Socio-Economic Index of Occupational Status   
-\end{itemize}
+\newpage
 %% If you have bibdatabase file and want bibtex to generate the bibitems, please use
@@
 \bibliography{Literature}
-In the working environment, health is not a trivial factor either. The \citet{who_global_1994} stresses the importance of health
+
 \end{document}

### 2019-01-31 LaTex/Paper_fulltext.tex
Added declaration of authorship

--- 2019-01-31 LaTex/Paper_fulltext.tex
@@
 Originally coming from environmental research this approach has already allowed researcher to survey wide ranges of locations and time periods and thus provides access to high-quality data that could quite impossibly be attained by professional research teams on their own \citep{bonney_citizen_2009}.
-
 	Conceptually, 'Open Science' and 'Citizen Science' are deeply intertwined as 'Citizen Science' needs to be as transparent and easily accessible as possible to be effective on a large scale \citep{nov_dusting_2011} and 'Open Science' needs to offer educational value and low entry barriers if it wants to gain attention and participation of a broader audience \citep{newman_future_2012}.
@@
 	\hline 
-	Geodata					&  		   				&  				 			\\ 
+	Geodata					&  		   X			&  				 			\\ 
 	\hline 
-	Git	     			 	&  		   				&  				 			\\ 
+	Git	     			 	&  		   				&  				X 			\\ 
 	\hline 			
-	Indices				 	&  		   				&  				 			\\ 
+	Indices				 	&  		   X			&  				 			\\ 
 	\hline 
@@
 	\hline 
-	Shiny 				 	&  		   				&  				 			\\ 
+	Shiny 				 	&  		   				&  				X 			\\ 
 	\hline 
-	Codebook 			 	&  		   				&  				 			\\ 
+	Codebook 			 	&  		   				&  				X 			\\ 
 	\hline 
-	Geoplots 			 	&  		   				&  				 			\\ 
+	Geoplots 			 	&  		   X			&  				 			\\ 
 	\hline 
@@
 	\hline 
-	Motivation			 	&  		   				&  				 			\\ 
+	Motivation			 	&  		   				&  				X 			\\ 
 	\hline 
-	Theory				 	&  		   				&  				 			\\ 
+	Theory				 	&  		   X			&  				 			\\ 
 	\hline 
@@
 
+	\newpage
+	\hangindent=\parindent
+	\parindent 0pt
+	\hangafter=-1
+	\textbf{Declaration of Authorship}
+	\\
+
+	\hangindent=\parindent
+	\parindent 0pt
+	\hangafter=-1
+	We hereby declare that the project paper and all related project elements submitted by the title 
+	\\
+
+	\hangindent=\parindent
+	\parindent 0pt
+	\hangafter=-1
+	\textbf{Visualising Life Satisfaction in Countries across Europe: Individual- and Multi-Level Evidence on Cultural and Pecuniary Factors} 
+	\\
+
+	\hangindent=\parindent
+	\parindent 0pt
+	\hangafter=-1
+	Are written unaided by ourselves as indicated in the section 'Responsibility for Project Areas'. 
+	\\
+
+	\hangindent=\parindent
+	\parindent 0pt
+	\hangafter=-1
+	All indirect or direct sources are acknowledged in the bibliography. We are aware that a false declaration is punishable by law. 
+	\\
+	\\
+	\\
+	Place, date                                         	Signature Noah Peters
+	\\
+	\\
+	\\
+	\\
+	\\
+	Place, date												Signature Nicolas Goeller
+
 \end{document}

### 2019-01-31 LaTex/Paper_fulltext.tex
Last bit of theory part

--- 2019-01-31 LaTex/Paper_fulltext.tex
@@
 
+\subsection{Macro-Level Variables and Multilevel Analyses}
+So far, the theoretical background sketched relations between life satisfaction and various individual-level variables. While
+this perspective enables us to assess psychological constructs and individual life realities in general, a core feature of our
+analysis is the incorporation of multilevel investigations. \\
+Being interested in life satisfaction across European countries and regions, we examined variables that might account for 
+between-variance. The background on these factors and their respective relation to life satisfaction is examined in this 
+section. We point out that the macro-variables cover a wider array than the work-related aspects described above. Nonetheless,
+the macro-variables and the multilevel models can be used to supplement individual-level analyses as it is common practice
+in multilevel contexts.
+
+\subsubsection{Social capital} 
+\citet{bartolini_happy_2014} inspect the relation between life satisfaction and social capital on a micro-level and resume that
+social capital can explain variation in life satisfaction in the medium- and long-run while the association is less striking in a
+short-run perspective. \citet{mikucka_when_2017} underpin this verdict with their multilevel analysis using World Values Study 
+and EVS data. While we did not obtain time-series data to inspect associations over time, we recommend to incorporate 
+social capital as a regional level-two variable (here: European NUTS1 regions). Referring to Putnam, \citet{Savelkoul_explaining_2011} imply that social 
+capital is especially related to the regional level. \\
+Having set the level of inspection, we focussed on the operationalisation of social capital given EVS data. \citet{van_schaik_social_2002}
+recommends a procedure specifically related to the EVS that builds on using several categories of social capital as a proxy
+for the overall concept. In this fashion, we constructed two indices for (i) institutional trust and (ii) trustworthiness that measures
+the prevalence of social norms legal compliance. Additionally, the EVS includes a binary item covering interpersonal trust. \citet{van_schaik_social_2002}
+also associates the respective categories of social capital with European NUTS1 regions, as we do. \\
+Comprehensive plots of the proxies can be found in our Observer app and multilevel regressions can be computed.
+
+\subsubsection{Economic macro-data}
+Building on \citet{easterlin_does_1974} and the diverse remarks on the association of life satisfaction (see above), GDP presents itself
+as a useful multilevel factor. We use country data on GDP for 2008 (the year of the EVS's fourth wave) as a level-3 variable.
+Furthermore, Human Development Index (HDI) data can be incorporated to capture a wider socio-economic context.
+
+\subsubsection{Freedom House Ratings}
+In order to inspect level-3 effects regarding differences in the political system, we offer to incorporate various Freedom House 
+indicators measuring a country's level of democracy. While this perspective offers an additional flavour, it lies not within the main
+focus of this project. \\
+A detailed description of all data used for this project is presented in the following section.
+
 \section{Data}
-All data employed in this project i acquired from official sources and with permission of the provider or in the course of open access.
+All data employed in this project is acquired from official sources and with permission of the provider or in the course of open access.
 Data as well as original codebooks will be made available for the user upon request, or delivered within the project.

### 2019-01-31 LaTex/Paper_fulltext.tex
Added user Manual

--- 2019-01-31 LaTex/Paper_fulltext.tex
@@
 \section{User Manual}
+The section is aimed at enabling the user to use the project results competently.
+To do so, a working model of the statistical software R is required, the following instructions are made for application within its GUI RStudio.
+The project is divided in 'Data preparation', "Varaible Preparation', 'Visualisation' and 'app'.
+The latter is the most important for the first time user as it will provide access to the reactive visualisation tool 'European Happiness Observer' that is the centrepiece of this project.
+First time users, may ignore the other files and can execute the whole 'app' file after installing all required packages.
+The visualisation tool will open itself subsequently, no further steps or data provision needed.
+The application has guidelines itself that should be helpful for the user in navigating through its function.
+Generally speaking, it offers opportunities to build mapplots for the nations and most of the regions at NUTS-1 level, other visualisations can be made using bar charts, bivariate line plots and boxplots grouped by geographical entities.
+Additionally, it is possible to adapt some linear regression models and display their output.
+Finally, raw data tables are displayed as the most unrefined view on the data.
+While the mapplots, bar charts and line plots can be understood quite intuitively, this paragraph is aimed at giving some guidelines for working with the 'Boxplot' and 'Regression table' elements in the application.\\
+
+A boxplot is a visualisation that displays the distribution of data on the y-axis,  The dark bar inside the box indicates the position of the median, while the box itself indicates the value for the first and third quartile of the data.
+Thus the inside of the box contains 50\% of the data.
+The antennas around the box (also called 'whiskers') have a distance from it of 1.5 times the size of the box (distance between the third and first quartile, called interquartile range).
+Any data points outside the range of box and whiskers are considered outliers from the data distribution.
+Thus, the boxplot allows to compare the spread of data for different groups that are displayed on the x-axis.
+In the application it is possible, to order the groups after certain variables to see, whether the variable may have an effect on the spread of data.
+
+\input{table1}
+
+The regression tables provide the user with the opportunity to compile customized regression models that will give formally statistical evidence on the substance of relations in the data.
+The simple linear regression without multilevel elements is to be interpreted very similar to the line plot, as starting with the constant value, every increase in the independent variable changes the dependent variable by a certain amount.
+The number of stars next to the independent variables indicate the significance of its effect (significance as in 'being present').
+The total validity of the model can be taken from the value of R$^2$ as it indicates the share of data spread in the dependent variable that is explained by the data spread in the independent variable.
+In table 1, this is exemplified for the model from 'Job satisfaction' on 'Life satisfaction'.\\
+
+\input{table2}
+
+However, by introducing multilevel elements into the regression model, the interpretation changes.
+Multilevel models rest on the assumption that the state of a variable as a characteristic of an individual (like life satisfaction) is expected to be influenced by structural factors.
+These influences can be accounted for when looking at the effect of variables that are measured on a structural level (like GDP per capita).
+The effect of this independent variable will then explain differences in the dependent variable that occur because there is variation between the groups, where the structural variable is measured (between-variance).
+It is then possible to account with variables for the individuals (like job satisfaction) exclusively for the variance inside the group (within-variance).
+This is exemplified in table 2.
+Like before, the significance of a variable is indicated by its number of stars.
+However, the quality of the model calculated is more difficult to assess than before.
+In table 2 the Akaike Information criterion (AIC) is provided for that purpose.
+This measure in absolute says nothing about the quality of the model but comparing several model, the one with the smallest value will be the one that explains the most.\\
+
+Users might be interested in seeing more detailed information on variables and can do so in the codebook 'EVS\_final\_cdbk' in the project file.
+It is also possible to generate the codebook anew by executing the 'Visualisations file'.
+if users may wish to make changes to data or transformation of variables, the may do so in the files 'Data preparation' and 'Variable preparation'.
 

### 2019-01-31 C_Analysis.R

@@ -6,53 +6,67 @@ library(tidyverse)
 library(lme4)
 library(sjstats)
 library(lmerTest)
-library(sjPlot)
+library(stargazer)
+
 
 #---------------------------------------------------------------------------------------
 
 #Read in dataset from B_Variable Preparation
 EVS_final <- read_rds("Data/EVS_final.rds")
 
-#Subsetting data and omitting NAs
-ana_dat <- select(EVS_final, nation, reg, sat, happ, job_sat, siops, isei, sup, entre, 
-                  inc)
-
-#----------------------------------------------------------------------------------------
-
-##Descriptive analysis
 
 #----------------------------------------------------------------------------------------
 
 ##  1. Individual model (Level 1)
 
-m1 <- lm(sat ~ job_sat + siops + isei, data = ana_dat)
+m1 <- lm(sat ~ job_sat, data = EVS_final)
 summary(m1)
+
+stargazer(m1,  #regression models 
+          type = "latex", # character vector (eg. "text" / "html" / "latex")
+          title = "Linear regression model",  # header
+          style = "default",  # style (choice see below)
+          summary = NULL,  # logical vector: output summary statistics when given data.frame
+          out = "LaTex/table1.tex", # path and output of file
+          out.header = FALSE, # logical vector: should output file contain code-header?
+          column.labels = c("Basic model"), # column labels for mod1/mod2
+          column.separate = c(1,1),  # how column labels should be assigned (label over sev. columns possible)
+          covariate.labels = c("Job satisfaction",
+                               "Constant"),
+          dep.var.caption = "Dependent Variable", # Caption (Top) of dependent variable
+          star.cutoffs = c(0.05,0.01,0.001),
+          dep.var.labels = c("Life satisfaction"))
 
 #---------------------------------------------------------------------------------------------
 
 ##  2. Regional models (Level 2)
 
-#Null model for regions
-m2 <- lmer(sat ~ 1 + (1|reg), data = ana_dat)
+m2 <- lme4::lmer(sat ~ job_sat + + (1|reg), data = ana_dat)
 summary(m2)
-icc(m2)
-
-#Model with Regional and Individual predictors
-m3 <- lmer(sat ~ 1 + (1|reg), data = ana_dat)
-summary(m3)
 
 #---------------------------------------------------------------------------------------------
 
 ##  3. National models (Level 3)
 
-#Null model for nations
-m4 <- lmer(sat ~ 1 + (1|nation), data = ana_dat)
-summary(m4)
-icc(m4)
[…]

### 2019-01-31 LaTex/Paper_fulltext.tex
Intro and Abstract

@@
 			%% Text of abstract
-			This project was set out to provide an easy access to a comprehensible way for exploring the differences in determinants of life satisfaction 
-			all over Europe. While the actual analysis and presentation of results will be shown in different ways, this document
-			shall provide an overview over the data used and the treatment of variables applied during the project. 
-			The project's aim is not only to inspire interest into the determinants of happiness in human life, but also to encourage
-			later researchers to critically enquire with the work done here in order to point out its shortcomings and improve the final
-			result. Thus this document should provide the necessary transparency about sourcing of data and work on variables that
-			has been done previously to conducting the analysis and visualising results.
+			This project was set out to provide a comprehensible way for exploring the eclectic determinants of life satisfaction 
+			across Europe. An interactive visualisation tool was developed to enable independent inspections of happiness-related
+			research questions. While the interactive tool covers the actual analysis and presentation of results, this document
+			provides an overview over data and variables used to contribute to contemporary happiness research. 
+			The project's aim is not only to spark interest on the determinants of happiness in human life, but also to encourage
+			later researchers to critically reflect upon the work presented here in order to point out potential shortcomings. 
+			Thus, this document should provide necessary transparency regarding data sourcing and variable selection.
+			Lastly, accounts regarding the relatedness of life satisfaction and work-related variables are presented to illustrate one potential
+			field of application of our interactive tool. 
 		\end{abstract}
@@
 
-	\section{Introduction and Preliminary remarks}
+	\section{Project Scope and Preliminary Remarks}
+	Our project investigates life satisfaction across Europe and puts an emphasis on the data visualisation part that comes
+	with communicating research results. This aspect appears most relevant with regard to `Open Science' and `Citizen
+	Science' approaches that are gaining importance. In order to cater to this motivation, we developed an interactive interface
+	- The European Happiness Observer - to inspect life satisfaction and its various antecedent across Europe. The app offers
+	a myriad of possibilities to inspect life satisfaction on a descriptive basis or delve deeper by computing regression and multilevel
+	analyses. Here, the user can select various options and combinations from a pre-selected pool of variables. \\
+	Conciously, the scope of our visualisation tool is kept wider and should steer curiosity and inclination to develop own 
+	happiness-related (research) questions. To assure at least some theoretical embeddedness, this paper elaborates upon the
+	relation of life satisfaction and work-related variables. This perspective served as a broad orientation during the development
+	phase of our tool, illustrates the variable selection and comprehensivly exhibits what kind of aspects happiness research
+	can elucidate. At this point it should be stressed that the Observer enables the user to examine the questions raised in this
+	paper. However, the tool's scope goes far beyond these applications and invites every user to get creative in her own 
+	right. Thus, this background paper fulfills two major functions: One the one hand, it communicates the data-driven
+	approach of our visualisations i.e., the European Happiness Observer app. On the other hand, it documents the theoretical foundation
+	of our variable selection and examplifies a major question of happiness research, namely the relation to work-related
+	variables.
+
+	\section{Motivational Manifesto}
+	In a time of widespread digitalisation for many aspects of modern society, 'datafication' and subsequent analysis of data has become ever more prevalent \citep{baack_datafication_2015, lycett_datafication:_2013}.
+Transparency of data sourcing, data access and methods employed in the course of this analysis is a chief concern of the 'Open Science' movement \citep{delfanti_open_2010}.
+With more and more transparency, reproducibility of procedures is expected to increase and thus the quality of insights by replication and and improvements of methods through a 'wisdom of the (educated) crowd'.
+There have been extensive developments in the form of guidelines and protocols for researchers, universities and scientific journals that are aimed at promoting more transparency in the scientific process \citep{nosek_promoting_2015, miguel_promoting_2014}.
+	At the same time, large shares of the population lack the essential skills to participate in the research process or perceive the barrier of entry in acquiring the necessary skills to do so, as insurmountably high.
+The concept of 'Citizen Science' has thus emerged sitting at the edge of both science and education.
+Citizen science projects enlist non-scientist members of the public to participate in the research (mainly in data collection), allowing to survey large amounts of data, while educating participants about the research topic \citep{shirk_public_2012}.
+Originally coming from environmental research this approach has already allowed researcher to survey wide ranges of locations and time periods and thus provides access to high-quality data that could quite impossibly be attained by professional research teams on their own \citep{bonney_citizen_2009}.
+	Conceptually, 'Open Science' and 'Citizen Science' are deeply intertwined as 'Citizen Science' needs to be as transparent and easily accessible as possible to be effective on a large scale \citep{nov_dusting_2011} and 'Open Science' needs to offer educational value and low entry barriers if it wants to gain attention and participation of a broader audience \citep{newman_future_2012}.
+We feel committed to the mission of these movements and have thus designed our project in order to advance it for the social sciences.
+
+	\section{Introduction to the Relation of Life Satisfaction and Work-Related Variables}
 	Anecdotal evidence suggests that people spend a significant share of their lifetime working (for a casual calculation 
@@
 	macroeconomic as well as developmental indicators were performed to account for variation across regions and countries.
-	%% Könntest du den Fokus ein wenig weg von Work und hin auf multilevel + pecuniary/ non-pecuiary setzen, für so viel Fokus haben wir nicht genug work-relatedness
-	\section{Motivational Manifesto}
-	In a time of widespread digitalisation for many aspects of modern society, 'datafication' and subsequent analysis of data has become ever more prevalent \citep{baack_datafication_2015, lycett_datafication:_2013}.
-Transparency of data sourcing, data access and methods employed in the course of this analysis is a chief concern of the 'Open Science' movement \citep{delfanti_open_2010}.
-With more and more transparency, reproducibility of procedures is expected to increase and thus the quality of insights by replication and and improvements of methods through a 'wisdom of the (educated) crowd'.
-There have been extensive developments in the form of guidelines and protocols for researchers, universities and scientific journals that are aimed at promoting more transparency in the scientific process \citep{nosek_promoting_2015, miguel_promoting_2014}.
-	At the same time, large shares of the population lack the essential skills to participate in the research process or perceive the barrier of entry in acquiring the necessary skills to do so, as insurmountably high.
-The concept of 'Citizen Science' has thus emerged sitting at the edge of both science and education.
-Citizen science projects enlist non-scientist members of the public to participate in the research (mainly in data collection), allowing to survey large amounts of data, while educating participants about the research topic \citep{shirk_public_2012}.
-Originally coming from environmental research this approach has already allowed researcher to survey wide ranges of locations and time periods and thus provides access to high-quality data that could quite impossibly be attained by professional research teams on their own \citep{bonney_citizen_2009}.
-	Conceptually, 'Open Science' and 'Citizen Science' are deeply intertwined as 'Citizen Science' needs to be as transparent and easily accessible as possible to be effective on a large scale \citep{nov_dusting_2011} and 'Open Science' needs to offer educational value and low entry barriers if it wants to gain attention and participation of a broader audience \citep{newman_future_2012}.
-We feel committed to the mission of these movements and have thus designed our project in order to advance it for the social sciences.
 
-	\section{Theory and Literature Review}
+	\section{Theoretical Background and Methodological Implementation}
 	Although the causal relation between work and happiness/life satisfaction seems straightforward and intuitively correct,

### 2019-01-31 LaTex/Paper_fulltext.tex
tables

--- 2019-01-31 LaTex/Paper_fulltext.tex
@@
 	macroeconomic as well as developmental indicators were performed to account for variation across regions and countries.
-	%% Könntest du den Fokus ein wenig weg von Work und hin auf multilevel + pecuniary/ non-pecuiary setzen, für so viel Fokus haben wir nicht genug work-relatedness
+
 	\section{Motivational Manifesto}
@@
 multiple factors that touch on leadership responsibilities, self-determination related factors, or the occupation's reputation.
-The following variables can be examined using the European Happiness Observer:
+The following variables can be examined using the European Happiness Observer among others:
 \begin{itemize}
@@
 Thus, the boxplot allows to compare the spread of data for different groups that are displayed on the x-axis.
-In the application it is possible, to order the groups after certain variables to see, whether the variable may have an effect on the spread of data.
+In the application it is possible, to order the groups after certain variables to see, whether the variable may have an effect on the spread of data.\\
 
@@
 The total validity of the model can be taken from the value of R$^2$ as it indicates the share of data spread in the dependent variable that is explained by the data spread in the independent variable.
-In table 1, this is exemplified for the model from 'Job satisfaction' on 'Life satisfaction'.\\
+In table 1, this is exemplified for the model from 'Job satisfaction' on 'Life satisfaction'.
 

### 2019-01-31 D_Visualisation.R
Changed lat/long for regional plots

@@ -128,7 +128,7 @@ ggplot(data = reg_geodata)+
   labs(fill = "Trustworthiness (from low to high)")+
   ggtitle("Regional Europe Map", 
           subtitle = "NUTS 1 Regions and Trustworthiness")+
-  coord_sf(xlim = c(-24, 32), ylim = c(33, 71), expand = FALSE)
+  coord_sf(xlim = c(-24, 50), ylim = c(33, 71), expand = FALSE)
 
 #     3.1.2 Regions and interpersonal trust
 ggplot(data = reg_geodata)+
@@ -136,7 +136,7 @@ ggplot(data = reg_geodata)+
   labs(fill = "Interpersonal trust (from low to high)")+
   ggtitle("Regional Europe Map", 
           subtitle = "NUTS 1 Regions and Interpersonal Trust")+
-  coord_sf(xlim = c(-24, 32), ylim = c(33, 71), expand = FALSE)
+  coord_sf(xlim = c(-24, 50), ylim = c(33, 71), expand = FALSE)
 
 #     3.1.2 Regions and institutional trust
 ggplot(data = reg_geodata)+
@@ -144,7 +144,7 @@ ggplot(data = reg_geodata)+
   labs(fill = "Institutional trust (from low to high)")+
   ggtitle("Regional Europe Map", 
           subtitle = "NUTS 1 Regions and Institutional Trust")+
-  coord_sf(xlim = c(-24, 32), ylim = c(33, 71), expand = FALSE)
+  coord_sf(xlim = c(-24, 50), ylim = c(33, 71), expand = FALSE)
 
 #     3.1.2 Regions and Perception of fairness
 ggplot(data = reg_geodata)+
@@ -152,7 +152,7 @@ ggplot(data = reg_geodata)+
   labs(fill = "Perception of fairness (from low to high)")+
   ggtitle("Regional Europe Map", 
           subtitle = "NUTS 1 Regions and Percieved  Fairness")+
-  coord_sf(xlim = c(-24, 32), ylim = c(33, 71), expand = FALSE)
+  coord_sf(xlim = c(-24, 50), ylim = c(33, 71), expand = FALSE)
 
 
 #----------------------------------------------------------------------------------------------------------

### 2019-01-31 LaTex/Paper_fulltext.tex
Geo data description

@@
 	vary across regions and countries? \\
-	Using data form the European Values Study (EVS) and several other sources, we calculated multiple regression models to assess the causality of
-	work-related factors and well-being on the individual level. Moreover, multi-level analyses with regard to social capital and
+	Using data form the European Values Study (EVS) and several other sources, we calculated multiple regression models to assess the causality of work-related factors and well-being on the individual level.
+Moreover, multi-level analyses with regard to social capital and
 	macroeconomic as well as developmental indicators were performed to account for variation across regions and countries.
@@
 The missing value of the HDI 2008 for Kosovo was approximated by the value for Serbia in 2008.
+\\
+To obtain national geo data for European countries, the `rnaturalearth' package was used \citep{south_rnaturalearth:_2017}.
+Geo data for European regions (NUTS1 2006) were obtained from Eurostat with the according package \citep{lahti_eurostat:_2019}. For the 
+processing of geo data, the `sf' package was employed \citep{pebesma_sf:_2018}.
 
@@
 	\hline 
-	Data wrangling			&  		   				&  				 			\\ 
+	Data wrangling			&  		   X			&  				X 			\\ 
 	\hline 
-	Variable preparation 	&  		   				&  				 			\\ 
+	Variable preparation 	&  		   X			&  				X 			\\ 
 	\hline 
@@
 	\hline 
-	Paper visualisations 	&  		   				&  				 			\\ 
+	Paper visualisations 	&  		   				&  				X 			\\ 
 	\hline 
@@
 	\hline 
-	LaTex				 	&  		   				&  				 			\\ 
+	LaTex				 	&  		   X			&  				 			\\ 
 	\hline 
@@
 	\hline 
-	User manual			 	&  		   				&  				 			\\ 
+	User manual			 	&  		   				&  				X 			\\ 
 	\hline 
@@
 	\\
-	\\
+	Copenhagen, January 31th 2019\\
 	\\
@@
 	Place, date												Signature Nicolas Goeller
+	\\
+	Lehrensteinsfeld, January 31th 2019
 

### 2019-01-31 LaTex/Paper_fulltext.tex
Added conclusion

@@
 	Science' approaches that are gaining importance. In order to cater to this motivation, we developed an interactive interface
-	- The European Happiness Observer - to inspect life satisfaction and its various antecedent across Europe. The app offers
+	- The European Happiness Observer - to inspect life satisfaction and its various antecedents across Europe. The app offers
 	a myriad of possibilities to inspect life satisfaction on a descriptive basis or delve deeper by computing regression and multilevel
 	analyses. Here, the user can select various options and combinations from a pre-selected pool of variables. \\
-	Conciously, the scope of our visualisation tool is kept wider and should steer curiosity and inclination to develop own 
+	Consciously, the scope of our visualisation tool is kept wider and should steer curiosity and inclination to develop own 
 	happiness-related (research) questions. To assure at least some theoretical embeddedness, this paper elaborates upon the
 	relation of life satisfaction and work-related variables. This perspective served as a broad orientation during the development
-	phase of our tool, illustrates the variable selection and comprehensivly exhibits what kind of aspects happiness research
+	phase of our tool, illustrates the variable selection and comprehensively exhibits what kind of aspects happiness research
 	can elucidate. At this point it should be stressed that the Observer enables the user to examine the questions raised in this
 	paper. However, the tool's scope goes far beyond these applications and invites every user to get creative in her own 
-	right. Thus, this background paper fulfills two major functions: One the one hand, it communicates the data-driven
+	right. Thus, this background paper fulfils two major functions: One the one hand, it communicates the data-driven
 	approach of our visualisations i.e., the European Happiness Observer app. On the other hand, it documents the theoretical foundation
-	of our variable selection and examplifies a major question of happiness research, namely the relation to work-related
+	of our variable selection and exemplifies a major question of happiness research, namely the relation to work-related
 	variables.
@@
 suggest that life satisfaction remains flat, despite increases in GDP per capita. These findings still heavily inform other researchers'
-work and indicate that the relative income in comparision to others might influence SWB \citep{di_tella_happiness_2010,drichoutis_reference_2010,pedersen_happiness_2011}.
+work and indicate that the relative income in comparison to others might influence SWB \citep{di_tella_happiness_2010,drichoutis_reference_2010,pedersen_happiness_2011}.
 Emphasising relative position follows the lead of behavioural labour economics that builds on the findings of prospect theory \citep{kahneman_prospect_1979}.
 This concept implies that reference points rather than absolute positions inform people's choices. In the course of behavioural
-labour economics, this backgroundis highly relevant \citep{dohmen_behavioral_2014}. 
+labour economics, this background is highly relevant \citep{dohmen_behavioral_2014}. 
 While investigations examining relative status and changes over time appear highly relevant and theoretically justified, the present
-project ist not suited to cover these questions. Not being in posession of either time series or relative income data, we need
+project is not suited to cover these questions. Not being in possession of either time series or relative income data, we need
 to focus on absolute income at a certain point in time. Although this independent variable might not capture the underlying relation
@@
 due to non-pecuniary circumstances that are innate and to the status of self-employment. \\
-This remark also leads to a greater discussion of features proposed by self-determinantion theory \citep{deci_intrinsic_1985,deci_self-determination_1989,ryan_self-determination_2000}.
+This remark also leads to a greater discussion of features proposed by self-determination theory \citep{deci_intrinsic_1985,deci_self-determination_1989,ryan_self-determination_2000}.
 Lastly, \citet{luo_essays_2018} uses EVS data estimate the effects of pecuniary and non-pecuniary factors on life satisfaction
@@
    \item Work importance in life
-   \item Freedem of job decisions
+   \item Freedom of job decisions
    \item Job development
@@
 The missing value of the HDI 2008 for Kosovo was approximated by the value for Serbia in 2008.
-\\
-To obtain national geo data for European countries, the `rnaturalearth' package was used \citep{south_rnaturalearth:_2017}.
-Geo data for European regions (NUTS1 2006) were obtained from Eurostat with the according package \citep{lahti_eurostat:_2019}. For the 
-processing of geo data, the `sf' package was employed \citep{pebesma_sf:_2018}.
 
@@
 \section{Conclusion}
+To conclude this project description it remains to say that this project was set out, trying to keep its stand regarding several different, sometimes conflicting expectations.
+The ambition to produce high quality analytical insights combining with deep-diving theoretical contextualisation that together contribute to the state of research in the discipline alone is a challenging one.
+It is met with an ambition of similar challenge in trying to produce a project that allows easy access exploration of the data in question for a lay public that has only little understanding of the statistics or programming involved, while at the same time offering enough substance to be of educational value for the user and inspire further interest for the topic and method of analysis.
+These are expectational challenges that are similar to the aforementioned trade-off between 'Open Science' and 'Citizen Science'.
+However, the result of this project is also promising in that it does fulfil many of the expectations that it was met with.
+The application produced is relatively easy accessible and offers a range of methods of analysis for a substantive amount of variables for individuals, regions and nations.
+It is possible to produce analytical results on many levels of scientific expertise and sourcing and processing of data is thoroughly documented.
+There is also a significant amount of theoretical consideration that flew in selection of variables and methods of analysis and that will hopefully inspire interest on th topic of life satisfaction.
+Among future challenges, there is the quality and access to data, especially for the analysis of countries and regions, which today is incomplete in many cases.
+There are some remaining challenges specifically for this project, if it is to achieve the its ultimate goal of offering access to a broad public audience, but we hope that with this preliminary project state there first important step are made with many more to come.
 

### 2019-01-31 European_Happiness_Observer/app.R
Final changes on the app

@@ -237,12 +237,12 @@ ui <- dashboardPage(
               )),
       tabItem(tabName = "regmap", h1("Geographical overview"),
               fluidRow(
-                box(title = "Europe map plot for regions", status = "primary", solidHeader = T, width = 9,
+                box(title = "Europe map plot for regions on NUTS-1 level", status = "primary", solidHeader = T, width = 9,
                     plotOutput("regiomap")),
                 box(title = "Controls for map plot", status = "warning", solidHeader = T, 
                     width = 3,
                     "Choose your variable for plotting",br(), br(),
-                    varSelectInput("regio", "Variable:", shiny_nat[,1:7]))
+                    varSelectInput("regio", "Variable:", shiny_reg[,c(3:6,8)]))
               )),
       
       tabItem(tabName = "graphics", h1("Graphics for detailed analysis"),
@@ -344,7 +344,21 @@ server <- function(input, output) {
     }
   })
   
-  output$regiomap <- renderPlot({})
+  output$regiomap <- renderPlot({
+    if(!!input$regio != "Country_of_residence" & !!input$regio != "Region_of_residence"){
+      ggplot(data = shiny_reg)+
+        geom_sf(aes(fill = !!input$regio))+
+        labs(fill = paste(input$regio))+
+        scale_fill_viridis_c()+
+        coord_sf(xlim = c(-24, 50), ylim = c(33, 71), expand = FALSE)
+    } else{
+      ggplot(data = shiny_reg)+
+        geom_sf(aes(fill = !!input$regio))+
+        labs(fill = paste(input$regio))+
+        coord_sf(xlim = c(-24, 50), ylim = c(33, 71), expand = FALSE)
+    }
+    
+  })
   
   output$cdb <- renderText(includeText("EVS_final_cdbk"))
   
@@ -600,7 +614,3 @@ shinyApp(ui, server)
 
 #End of app code##------------------------------------------------------------------------------------
 #------------------------------------------------------------------------------------------
-EVS_final %>% 
-  select(reg) %>% 
-  summarise(mean())
-mean(EVS_final$reg, na.rm = T)

### 2019-01-31 LaTex/Paper_fulltext.tex
Altered conclusion

@@
 \section{Conclusion}
-To conclude this project description it remains to say that this project was set out, trying to keep its stand regarding several different, sometimes conflicting expectations.
-The ambition to produce high quality analytical insights combining with deep-diving theoretical contextualisation that together contribute to the state of research in the discipline alone is a challenging one.
-It is met with an ambition of similar challenge in trying to produce a project that allows easy access exploration of the data in question for a lay public that has only little understanding of the statistics or programming involved, while at the same time offering enough substance to be of educational value for the user and inspire further interest for the topic and method of analysis.
-These are expectational challenges that are similar to the aforementioned trade-off between 'Open Science' and 'Citizen Science'.
-However, the result of this project is also promising in that it does fulfil many of the expectations that it was met with.
-The application produced is relatively easy accessible and offers a range of methods of analysis for a substantive amount of variables for individuals, regions and nations.
-It is possible to produce analytical results on many levels of scientific expertise and sourcing and processing of data is thoroughly documented.
-There is also a significant amount of theoretical consideration that flew in selection of variables and methods of analysis and that will hopefully inspire interest on th topic of life satisfaction.
-Among future challenges, there is the quality and access to data, especially for the analysis of countries and regions, which today is incomplete in many cases.
-There are some remaining challenges specifically for this project, if it is to achieve the its ultimate goal of offering access to a broad public audience, but we hope that with this preliminary project state there first important step are made with many more to come.
-
+To conclude this project description, it remains to say that the project outline was, at times, torn between several 
+different, sometimes conflicting expectations. The ambition to produce high-quality analytical insights combined with 
+deep-diving theoretical contextualisation is a challenging endeavour. Conducting a project that allows easily accessible 
+data exploration for lay people is tough in its own right. One the one hand, one has to cater to an audience that has only
+little understanding of the statistics or programming behing the figures. On the other hand, 
+offering enough substance and educational value to spark further interest on the topic and the method should be assured.
+These are expectational challenges that are similar to the aforementioned trade-off between 'Open Science' and 'Citizen
+Science'. However, the result of this project is promising in so far as it fulfils many of the expectations that it was set out
+to meet. The application produced is easily accessible and offers a range of possibilities for the analysis of various
+variables on an individuals, regional, and national level. 
+Moreover, data sourcing and processing is thoroughly documented. Lastly, significant theoretical considerations
+justify our selection of variables and methods and will hopefully attract others to explore the topic of life satisfaction.\\
+Future challenges include sufficient data access and quality, especially regarding countries and regions that are not covered to date.
+Despite - or maybe because - of future challenges we are optimistic that this project is just the first destination of a longer journey. 
 	%% The Appendices part is started with the command \appendix; appendix sections are then done as normal sections

### 2019-01-31 LaTex/Paper_fulltext.tex
Proof-reading

@@
 Transparency of data sourcing, data access and methods employed in the course of this analysis is a chief concern of the 'Open Science' movement \citep{delfanti_open_2010}.
-With more and more transparency, reproducibility of procedures is expected to increase and thus the quality of insights by replication and and improvements of methods through a 'wisdom of the (educated) crowd'.
+With more and more transparency, reproducibility of procedures is expected to increase and thus the quality of insights by replication and improvements of methods through a 'wisdom of the (educated) crowd'.
 There have been extensive developments in the form of guidelines and protocols for researchers, universities and scientific journals that are aimed at promoting more transparency in the scientific process \citep{nosek_promoting_2015, miguel_promoting_2014}.
@@
 	happiness, comes in a four-point scale. Consequently, we only included the life satisfaction variable for our regression
-	models. For the remains of this study, we concentrate on life satisfaction as a dependent variable. The variable selection
-	is further elaborated upon in the method section.
+	models. For the remains of this study, we concentrate on life satisfaction as a dependent variable.
 
@@
 	meta-analysis, show a positive relationship between all three respective components of SWB and job satisfaction.
-	Going even further, \citet{near_job_1987} postulate that “job satisfaction must logically constitute some part of the
+	Going even further, \citet{near_job_1987} postulate that 
+	\begin{quotation}
+	“job satisfaction must logically constitute some part of the
 	broader construct of life satisfaction; thus, the question regarding spillover between the two must be ‘how much’ not
-	‘whether’." (p. 398). \\
-	Before we examine the how much-question, it is necessary to discuss the direction of causality between SWB and job
+	‘whether’." (p. 398). 
+	\end{quotation}
+
+	Before we examine the `how much-question', it is necessary to discuss the direction of causality between SWB and job
 	satisfaction. In general, the present literature differentiates two approaches: the part-whole theory
@@
 	above: job satisfaction constitutes a sub-category of overall life satisfaction \citep[cf.]{near_job_1987}.
-	Consequently, the approach is also conceived of as bottom-up approach \citep{erdogan_whistle_2012}. Conversely, the
+	Consequently, the approach is also conceived of as a bottom-up approach \citep{erdogan_whistle_2012}. Conversely, the
 	dispositional approach assumes a general predisposition of satisfaction with life to affect more specific domains like work
@@
 	variables that characterise the working environment beyond perceived job satisfaction. This wider scope appears
-	reasonable since “job satisfaction is only a partial indicator of satisfaction with work domain. A person’s satisfaction with
+	reasonable since 
+	\begin{quotation}
+	“job satisfaction is only a partial indicator of satisfaction with work domain. A person’s satisfaction with
 	career and work-related stress are important elements of the work domain that also need to be understood.” \citep[p. 1042]{erdogan_whistle_2012}.
+	\end{quotation}
 
@@
 In order to justify our variable selection the following remarks elucidate the respective variables' interplay with life satisfaction.
-In contrast to the previous examination of the overall concept of SWB, to accounts on the different variables are presented 
+In contrast to the previous examination of the overall concept of SWB, two accounts on the different variables are presented 
 in a briefer manner. Rather than covering the entire theoretical background, the following elaboration is though to spur interest
@@
 Self-employment as such exhibits a variable that is thought to affect life satisfaction \citep{andersson_happiness_2008,schneck_why_2014,van_der_zwan_self-employment_2018}
-due to non-pecuniary circumstances that are innate and to the status of self-employment. \\
-This remark also leads to a greater discussion of features proposed by self-determination theory \citep{deci_intrinsic_1985,deci_self-determination_1989,ryan_self-determination_2000}.
-Lastly, \citet{luo_essays_2018} uses EVS data estimate the effects of pecuniary and non-pecuniary factors on life satisfaction
-The EVS, for example, captures to which degree respondents feel the need for a job that enables employees to develop their talents. \\
+due to non-pecuniary circumstances that are innate to the status of self-employment. \\
+This remark also leads to a greater discussion of features proposed by self-determination theory \citep{deci_intrinsic_1985,deci_self-determination_1989,ryan_self-determination_2000}. \\
+Lastly, \citet{luo_essays_2018} uses EVS data to estimate the effects of pecuniary and non-pecuniary factors on life satisfaction
 of the unemployed. Intuitively, financial uncertainty might appear as the major concern of unemployment but further investigations
@@
 \subsubsection{Non-pecuniary factors of the work environment}
-Drawing on the previous remarks regarding in incorporation of work-related non-pecuniary variables, our analysis comprises of
+Drawing on the previous remarks regarding the incorporation of work-related non-pecuniary variables, our analysis comprises of
 multiple factors that touch on leadership responsibilities, self-determination related factors, or the occupation's reputation.
-The following variables can be examined using the European Happiness Observer among others:
+The following variables (amongst others) can be examined using the European Happiness Observer:
 \begin{itemize}
@@
 for the overall concept. In this fashion, we constructed two indices for (i) institutional trust and (ii) trustworthiness that measures
-the prevalence of social norms legal compliance. Additionally, the EVS includes a binary item covering interpersonal trust. \citet{van_schaik_social_2002}
+the prevalence of social norms and legal compliance. Additionally, the EVS includes a binary item covering interpersonal trust. \citet{van_schaik_social_2002}
 also associates the respective categories of social capital with European NUTS1 regions, as we do. \\
@@
 may be settled with the publisher or from the original codebook.
-The file contains a total 66281 observations for 477 variables over 46 European countries.
+The file contains a total of 66281 observations for 477 variables over 46 European countries.
 All transformations on variables can be read in detail from the codebook.

### 2019-01-31 LaTex/Paper_fulltext.tex
Final ones

--- 2019-01-31 LaTex/Paper_fulltext.tex
@@
 	analyses. Here, the user can select various options and combinations from a pre-selected pool of variables. \\
-	Consciously, the scope of our visualisation tool is kept wider and should steer curiosity and inclination to develop own 
+	Consciously, the scope of our visualisation tool is kept wide and should steer curiosity and inclination to develop own 
 	happiness-related (research) questions. To assure at least some theoretical embeddedness, this paper elaborates upon the
@@
 	its academic investigation is far from trivial. On the one hand, this appears in light of definitional questions regarding the 
-	wo variables of interest. On the other hand, operationalising the constructs appropriately remains subject to discussions
+	two variables of interest. On the other hand, operationalising the constructs appropriately remains subject to discussions
 	(for subjective well-being see: \citet{kahneman_developments_2006,layard_measuring_2010}). Lastly, the causal nature of job and life
@@
 	‘whether’." (p. 398). \\
-	Before we examine the how much-question, it is necessary to discuss the direction of causality between SWB and job
+	Before examining the how much-question, it is necessary to discuss the direction of causality between SWB and job
 	satisfaction. In general, the present literature differentiates two approaches: the part-whole theory
@@
 In contrast to the previous examination of the overall concept of SWB, to accounts on the different variables are presented 
-in a briefer manner. Rather than covering the entire theoretical background, the following elaboration is though to spur interest
-and curiosity to examine the relation of life satisfaction and other factors on one's own - aided by our interactive 
+in a briefer manner. Rather than covering the entire theoretical background, the following elaboration is thought to spur interest
+and curiosity in examining the relation of life satisfaction and other factors on one's own - aided by our interactive 
 European Happiness Observer interface.
@@
 To do so, a working model of the statistical software R is required, the following instructions are made for application within its GUI RStudio.
-The project is divided in 'Data preparation', "Varaible Preparation', 'Visualisation' and 'app'.
-The latter is the most important for the first time user as it will provide access to the reactive visualisation tool 'European Happiness Observer' that is the centrepiece of this project.
-First time users, may ignore the other files and can execute the whole 'app' file after installing all required packages.
+The project is divided in 'Data preparation', "Variable Preparation', 'Analysis, 'Visualisation' and 'app'.
+The latter is the most important for the first-time user as it will provide access to the reactive visualisation tool 'European Happiness Observer' that is the centrepiece of this project.
+First time users may ignore the other files and can execute the whole 'app' file after installing all required packages.
 The visualisation tool will open itself subsequently, no further steps or data provision needed.
-The application has guidelines itself that should be helpful for the user in navigating through its function.
-Generally speaking, it offers opportunities to build mapplots for the nations and most of the regions at NUTS-1 level, other visualisations can be made using bar charts, bivariate line plots and boxplots grouped by geographical entities.
-Additionally, it is possible to adapt some linear regression models and display their output.
+The application has guidelines itself that should be helpful for the user in navigating through its functions.
+Generally speaking, the app offers opportunities to build mapplots for the nations and most of the regions at NUTS-1 level, other visualisations can be made using bar charts, bivariate line plots and boxplots grouped by geographical entities.
+Additionally, it is possible to adapt some linear and multilevel regression models and display their output.
 Finally, raw data tables are displayed as the most unrefined view on the data.
@@
 
-A boxplot is a visualisation that displays the distribution of data on the y-axis,  The dark bar inside the box indicates the position of the median, while the box itself indicates the value for the first and third quartile of the data.
+A boxplot is a visualisation that displays the distribution of data on the y-axis, the dark bar inside the box indicates the position of the median, while the box itself indicates the value for the first and third quartile of the data.
 Thus the inside of the box contains 50\% of the data.
@@
 
-The regression tables provide the user with the opportunity to compile customized regression models that will give formally statistical evidence on the substance of relations in the data.
+The regression tables provide the user with the opportunity to compile customized regression models that will give formal statistical evidence on the substance of relations in the data.
 The simple linear regression without multilevel elements is to be interpreted very similar to the line plot, as starting with the constant value, every increase in the independent variable changes the dependent variable by a certain amount.
@@
 Users might be interested in seeing more detailed information on variables and can do so in the codebook 'EVS\_final\_cdbk' in the project file.
-It is also possible to generate the codebook anew by executing the 'Visualisations file'.
-if users may wish to make changes to data or transformation of variables, the may do so in the files 'Data preparation' and 'Variable preparation'.
+It is also possible to generate the codebook anew by executing the 'Visualisation' file.
+If users wish to make changes to data or transformation of variables, the may do so in the files 'Data preparation' and 'Variable preparation'.
 
@@
 different, sometimes conflicting expectations. The ambition to produce high-quality analytical insights combined with 
-deep-diving theoretical contextualisation is a challenging endeavour. Conducting a project that allows easily accessible 
+deep-diving theoretical contextualisation is a challenging endeavour. But also conducting a project that allows easily accessible 
 data exploration for lay people is tough in its own right. One the one hand, one has to cater to an audience that has only

### 2019-01-31 LaTex/Paper_fulltext.tex

@@
 	meta-analysis, show a positive relationship between all three respective components of SWB and job satisfaction.
-	Going even further, \citet{near_job_1987} postulate that “job satisfaction must logically constitute some part of the
+	Going even further, \citet{near_job_1987} postulate that 
+	\begin{quotation}
+	“job satisfaction must logically constitute some part of the
 	broader construct of life satisfaction; thus, the question regarding spillover between the two must be ‘how much’ not
-	‘whether’." (p. 398). \\
-	Before examining the how much-question, it is necessary to discuss the direction of causality between SWB and job
+	‘whether’." (p. 398). 
+	\end{quotation}
+
+	Before we examine the `how much-question', it is necessary to discuss the direction of causality between SWB and job
 	satisfaction. In general, the present literature differentiates two approaches: the part-whole theory
